---
title: "Reproducible Research: Peer Assessment 1"
author: "skydome20" 
date: "2015-06-05, Friday"
output:
  html_document:
    css: style.css
    
---


```{r setOpetions, echo=FALSE, results='hide'}
Sys.setlocale("LC_TIME", "English")
```

## 1.Loading and preprocessing the data
```{r, results='hold'}    
activity <- read.csv("activity.csv")
head(activity)
```

## 2.What is mean total number of steps taken per day?
Here I call function ```aggregate``` to count steps by per day
```{r}
steps_per_day <- aggregate(activity$steps, by=list(activity$date), FUN=sum, na.rm=TRUE) #main code
names(steps_per_day) <- c('date', 'steps')  #assign column names
head(steps_per_day)
```


Then create a histogram :
```{r}
barplot(steps_per_day$steps, names.arg=steps_per_day$date)

```


Finally, get the mean and median of steps taken per day:
```{r}
mean(steps_per_day$steps)
median(steps_per_day$steps)
```

## 3.What is the average daily activity pattern?
Also, use ```aggregate``` function to achieve the goal of averaging steps by time series
```{r}
means_per_interval <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE) # main code
names(means_per_interval) <- c("interval", "mean_of_steps") # assign column names
head(means_per_interval)
```


Then draw a line chart with x="time interval" and y="averaged steps"
```{r}
plot(x=means_per_interval$interval, y=means_per_interval$mean_of_steps, type="l", col="blue", xlab="interval", ylab="mean_of_steps")
```


And the max averaged steps is:
```{r}
max <- which.max(means_per_interval$mean_of_steps) # find the max steps, and return row_number/index
means_per_interval[max, "interval"]
```


## 4.Imputing missing values
```{r}
sum( is.na(activity$steps)==TRUE )
```


The strategy I use to fill missing values is to fill NAs with "mean for 5-minute interval"(M5MI)
Thus, first I create a new dataset:
```{r, results='hide'}
new.activity <- read.csv("activity.csv")
```


Second, find which row has NAs(missing value) on the "steps" column:
```{r}
na_rows <- which(is.na(new.activity$steps) )
head(na_rows)   #row_number indicates where is NA
```


Finally, assign M5MI(mean for 5-minute interval) to NA
```{r, results='hide'}
for (i in 1:length(na_rows)){
    index <- which( means_per_interval$interval == new.activity[na_rows[i], "interval" ] )
    new.activity[na_rows[i], "steps"] <- means_per_interval[index, "mean_of_steps"]
}
```
```{r}
table(is.na(new.activity$steps))
```



Now, plot a historgam again to compared with the previous hist and results:

It seems different: some days originally have nothing but now are filled. 
```{r}
new.steps_per_day <- aggregate(new.activity$steps, by=list(new.activity$date), FUN=sum) #count steps by date
names(new.steps_per_day) <- c('date', 'steps')  #assign column names
par(mfrow=c(1,2))
#old
barplot(steps_per_day$steps, names.arg=steps_per_day$date, main="old")
#new
barplot(new.steps_per_day$steps, names.arg=new.steps_per_day$date, main="new")  
``` 


And the mean and median:
```{r}
old_mean <-   mean(steps_per_day$steps)
old_median <- median(steps_per_day$steps)
new_mean <-   mean(new.steps_per_day$steps)
new_median <- median(new.steps_per_day$steps)

c(old_mean, new_mean)
c(old_median, new_median)
```

Since we replaces NAs with M5MI, the mean and median are definitely changed! 


## 5.Are there differences in activity patterns between weekdays and weekends?
Create a new varable 'weeklevel' contains 2 factors("weekday", "weekend")
```{r}
new.activity$date <- as.Date(new.activity$date, "%Y-%m-%d")
    
weeks <- weekdays(new.activity$date)
daylevel <- vector()

for (i in 1:length(weeks)){
    if( (weeks[i]=="Sunday") | (weeks[i]=="Saturday") ){
        daylevel[i] <- "weekend"
    }
    else{
        daylevel[i] <- "weekday"
    }
}

```
```{r}
new.activity$weeklevel <- factor(daylevel)
head(new.activity)
```


Similarly, use ```aggregate``` function to calculate means, but this time by 'interval' and 'weeklevel' 
```{r}
meanSteps_by_interval_week <- aggregate(new.activity$steps, by=list(new.activity$interval, new.activity$weeklevel), FUN=mean) 
names(meanSteps_by_interval_week) <- c("interval", "weeklevel", "steps") #assign column names
```


In the end, use Lattice plotting system to draw a line chart 
```{r}
library(lattice)
xyplot(steps ~ interval | weeklevel, meanSteps_by_interval_week, layout = c(1, 2), type = "l", xlab="Interval", ylab="Number of steps")
```



